---
title: "Plot_gone"
output: html_document
date: "2024-07-01"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('~/Desktop/Rworks')
```

## GONE Ne estimates

Plot Mediterranean samples (obsolete)

```{r}
setwd('~/Desktop/Rworks/')

df <- read.table('Output_Ne_Med.chr20.GQ20.DP8.edited',skip = 1,header = T)
df2 <- read.table('Output_Ne_Med.chr20.GQ20.DP8.edited_1',skip = 1,header = T)

row_num <- min(nrow(df),nrow(df2))
df_Med_chr20 <- merge(df[1:row_num,],df2[1:row_num,],by.x = 'Generation',by.y = 'Generation')
#The first ~200 generations are accurate with this method
library(ggplot2)

#plot 2 replicates in a single graph
png(filename = 'GONE_Med_chr20(2x).png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(df_Med_chr20[1:200,], aes(x = Generation)) +
  geom_line(aes(y = Geometric_mean.x), color = 'red') +
  geom_line(aes(y = Geometric_mean.y), color = 'blue') +
  ggtitle(paste('Demographic trajectory of Mediterranean population estimated by GONE, chr20')) + #add a title
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
dev.off()
```

### Plot Mallorca samples

```{r}
df_Mal <- read.table('GONE_output/Output_Ne_Mal.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)

#just plot the curve without including replicates
png(filename = 'GONE_Mall_1.2M_0.01_500gens+200rep.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(df_Mal[1:500,], aes(x = Generation)) +
  geom_line(aes(y = Geometric_mean), color = 'red') +
  ggtitle(paste('Demographic trajectory of Mediterranean population estimated by GONE')) + #add a title
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
dev.off()



#generate a list to store the internal replicates
file_list <- list.files(path = "GONE_output/TEMP/TEMP_Mal_0.01_rep200/", pattern = "*", full.names = TRUE)
df_ind <- list()
for (file in file_list) {
  # Read the TXT file
  data <- read.table(file, skip = 1,header = TRUE)
  # Extract the file name without extension to use as list name
  file_name <- tools::file_path_sans_ext(basename(file))
  # Store the data in the list with the file name as the list name
  df_ind[[file_name]] <- data
}
# Print the names of the dataframes in the list
print(names(df_ind))

#get the minimum row number for all replicates and generate a summary table
row_numbers <- unlist(lapply(df_ind, nrow))
n <- length(df_ind)
row_num_min <- min(c(nrow(df_Mal),row_numbers))
for (i in 1:n) {
  df_Mal <- cbind(df_Mal[1:row_num_min,],df_ind[[i]][1:row_num_min,2])
}

colname_reps <- paste('rep',seq(1,n))
colnames(df_Mal) <- c(colnames(df_Mal)[1:2],colname_reps)


library(tidyr)
library(ggplot2)


# Reshape data from wide to long format
test <- gather(df_Mal[21:500,], key = "column", value = "value", -Generation)
colors <- rep('red',n+1)  # Red and lighter red
names(colors) <- unique(test$column)  # Assign colors to levels
colors[-1] <- adjustcolor(colors[-1], alpha.f = 0.1) #Lighter colours for the internal replicates


png(filename = 'GONE_Mall_1.2M_0.01_21-500gens+200rep(CI).png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(test, aes(x = Generation, y = value, color = column)) +
  geom_line() +  # Overall alpha for all lines
  scale_color_manual(values = colors) +
  scale_y_continuous(limits = c(0, 5e4)) +
  scale_x_continuous(limits = c(20, 501)) +
  labs(x = "Generations", y = "Ne estimate", title = "") +
  theme(legend.position = "none")
dev.off()


```

If we exclude the replicates with 0s in the prediction:

```{r}
df_Mal <- read.table('GONE_output/Output_Ne_Mal.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)




#generate a list to store the internal replicates
file_list <- list.files(path = "GONE_output/TEMP/TEMP_Mal_0.01_rep200/", pattern = "*", full.names = TRUE)
df_ind <- list()
for (file in file_list) {
  # Read the TXT file
  data <- read.table(file, skip = 1,header = TRUE)
  # Extract the file name without extension to use as list name
  file_name <- tools::file_path_sans_ext(basename(file))
  # Store the data in the list with the file name as the list name
  df_ind[[file_name]] <- data
}
# Print the names of the dataframes in the list
print(names(df_ind))

#get the minimum row number for all replicates and generate a summary table
row_numbers <- unlist(lapply(df_ind, nrow))
n <- length(df_ind)
row_num_min <- min(c(nrow(df_Mal),row_numbers))
for (i in 1:n) {
  df_Mal <- cbind(df_Mal[1:row_num_min,],df_ind[[i]][1:row_num_min,2])
}

colname_reps <- paste('rep',seq(1,n))
colnames(df_Mal) <- c(colnames(df_Mal)[1:2],colname_reps)

#exclude small estimates and recalculate the GONE average
df_Mal_nofails <- cbind(df_Mal[,1],df_Mal[,df_Mal[nrow(df_Mal),]>1000])
colnames(df_Mal_nofails)[1] <- 'Generation'
geometric_mean <- function(x) {
  exp(mean(log(x)))
}
df_Mal_nofails$Geometric_mean <-  apply(df_Mal_nofails[,-1], 1, geometric_mean)

library(tidyr)
library(ggplot2)


# Reshape data from wide to long format
test <- gather(df_Mal_nofails[21:500,], key = "column", value = "value", -Generation)
colors <- rep('red',n+1)  # Red and lighter red
names(colors) <- unique(test$column)  # Assign colors to levels
colors[-1] <- adjustcolor(colors[-1], alpha.f = 0.1) #Lighter colours for the internal replicates


png(filename = 'GONE_Mall_1.2M_0.01_21-500gens+200rep(CI)_nofails.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(test, aes(x = Generation, y = value, color = column)) +
  geom_line() +  # Overall alpha for all lines
  scale_color_manual(values = colors) +
  scale_y_continuous(limits = c(0, 5e4)) +
  scale_x_continuous(limits = c(20, 501)) +
  labs(x = "Generations", y = "Ne estimate", title = "") +
  theme(legend.position = "none")
dev.off()

```

### Plot African samples with 1.2M SNPs across the autosome

```{r}
df_Afr <- read.table('GONE_output/Output_Ne_Afr.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_edited',skip = 1,header = T)


#generate a list to store the internal replicates
file_list <- list.files(path = "GONE_output/TEMP/TEMP_Afr/", pattern = "*", full.names = TRUE)
df_ind <- list()
for (file in file_list) {
  # Read the TXT file
  data <- read.table(file, skip = 1,header = TRUE)
  # Extract the file name without extension to use as list name
  file_name <- tools::file_path_sans_ext(basename(file))
  # Store the data in the list with the file name as the list name
  df_ind[[file_name]] <- data
}
# Print the names of the dataframes in the list
print(names(df_ind))

#get the minimum row number for all replicates and generate a summary table
row_numbers <- unlist(lapply(df_ind, nrow))
n <- length(df_ind)
row_num_min <- min(c(nrow(df_Afr),row_numbers))
for (i in 1:n) {
  df_Afr <- cbind(df_Afr[1:row_num_min,],df_ind[[i]][1:row_num_min,2])
}

colname_reps <- paste('rep',seq(1,n))
colnames(df_Afr) <- c(colnames(df_Afr)[1:2],colname_reps)


library(tidyr)
library(ggplot2)


#Reshape data from wide to long format
test <- gather(df_Afr[20:500,], key = "column", value = "value", -Generation)
colors <- rep('red',n+1)  # Red and lighter red
names(colors) <- unique(test$column)  # Assign colors to levels
colors[-1] <- adjustcolor(colors[-1], alpha.f = 0.1) #Lighter colours for the internal replicates

png(filename = 'GONE_Afr_1.2M_20-500gens+40rep.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(test, aes(x = Generation, y = value, color = column)) +
  geom_line() + 
  #scale_y_continuous(limits = c(0, 5e4)) +
  scale_x_continuous(limits = c(20, 501)) +
  scale_color_manual(values = colors) +
  labs(x = "Generations", y = "Ne estimate", title = "") +
  theme(legend.position = "none")
dev.off()
```

### Plot Mediterranean population excluding the Mallorca island group

```{r}
df_subMed <- read.table('GONE_output/Output_Ne_subMed.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)

library(ggplot2)
#just plot the curve without including replicates
png(filename = 'GONE_subMed_1.2M_500gens+200rep.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(df_subMed[1:500,], aes(x = Generation)) +
  geom_line(aes(y = Geometric_mean), color = 'red') +
  ggtitle(paste('Demographic trajectory of subMed group estimated by GONE')) + #add a title
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
dev.off()




file_list <- list.files(path = "GONE_output/TEMP/TEMP_subMed_rep200/", pattern = "*", full.names = TRUE)
df_ind <- list()
for (file in file_list) {
  # Read the TXT file
  data <- read.table(file, skip = 1,header = TRUE)
  # Extract the file name without extension to use as list name
  file_name <- tools::file_path_sans_ext(basename(file))
  # Store the data in the list with the file name as the list name
  df_ind[[file_name]] <- data
}
# Print the names of the dataframes in the list
print(names(df_ind))

#get the minimum row number for all replicates and generate a summary table
row_numbers <- unlist(lapply(df_ind, nrow))
n <- length(df_ind)
row_num_min <- min(c(nrow(df_subMed),row_numbers))
for (i in 1:n) {
  df_subMed <- cbind(df_subMed[1:row_num_min,],df_ind[[i]][1:row_num_min,2])
}

colname_reps <- paste('rep',seq(1,n))
colnames(df_subMed) <- c(colnames(df_subMed)[1:2],colname_reps)


library(tidyr)
library(ggplot2)


# Reshape data from wide to long format
test <- gather(df_subMed[20:500,], key = "column", value = "value", -Generation)
colors <- rep('red',n+1)  # Red and lighter red
names(colors) <- unique(test$column)  # Assign colors to levels
colors[-1] <- adjustcolor(colors[-1], alpha.f = 0.05) #Lighter colours for the internal replicates


png(filename = 'GONE_subMed_1.2M_20-500gens+200rep.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(test, aes(x = Generation, y = value, color = column)) +
  geom_line() +  # Overall alpha for all lines
  scale_color_manual(values = colors) +
  #ylim(0,5.5e04) +  
  scale_y_continuous(limits = c(0, 5.5e4)) +
  scale_x_continuous(limits = c(20, 501)) +
  labs(x = "Generations", y = "Ne estimate", title = "") +
  theme(legend.position = "none")
dev.off()
```

### Plot two groups in Mediterranean region together

```{r}
df_Mal <- read.table('GONE_output/Output_Ne_Mal.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)
df_subMed <- read.table('GONE_output/Output_Ne_subMed.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)


library(ggplot2)
#just plot the curve without including replicates
png(filename = 'GONE_2groups_1.2M_20-500gens+200rep.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(df_subMed[20:500,], aes(x = Generation)) +
  geom_line(aes(y = Geometric_mean), color = 'red') +
  ggtitle(paste('Combined demographic trajectories of two groups from Mediterranean region')) + #add a title
  geom_line(data = df_Mal[20:500,], aes(x = Generation, y = Geometric_mean), color = 'blue') +
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
dev.off()
```

If we exclude all the failed attempts within the Mallorca samples

```{r}
df_Mal <- read.table('GONE_output/Output_Ne_Mal.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)

#generate a list to store the internal replicates
file_list <- list.files(path = "GONE_output/TEMP/TEMP_Mal_0.01_rep200/", pattern = "*", full.names = TRUE)
df_ind <- list()
for (file in file_list) {
  # Read the TXT file
  data <- read.table(file, skip = 1,header = TRUE)
  # Extract the file name without extension to use as list name
  file_name <- tools::file_path_sans_ext(basename(file))
  # Store the data in the list with the file name as the list name
  df_ind[[file_name]] <- data
}
# Print the names of the dataframes in the list
print(names(df_ind))

#get the minimum row number for all replicates and generate a summary table
row_numbers <- unlist(lapply(df_ind, nrow))
n <- length(df_ind)
row_num_min <- min(c(nrow(df_Mal),row_numbers))
for (i in 1:n) {
  df_Mal <- cbind(df_Mal[1:row_num_min,],df_ind[[i]][1:row_num_min,2])
}

colname_reps <- paste('rep',seq(1,n))
colnames(df_Mal) <- c(colnames(df_Mal)[1:2],colname_reps)

#exclude small estimates and recalculate the GONE average
df_Mal_nofails <- cbind(df_Mal[,1],df_Mal[,df_Mal[nrow(df_Mal),]>1000])
colnames(df_Mal_nofails)[1] <- 'Generation'
geometric_mean <- function(x) {
  exp(mean(log(x)))
}
df_Mal_nofails$Geometric_mean <-  apply(df_Mal_nofails[,-1], 1, geometric_mean)


df_subMed <- read.table('GONE_output/Output_Ne_subMed.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.1.2M_rep200_edited',skip = 1,header = T)


library(ggplot2)
#just plot the curve without including replicates
png(filename = 'GONE_2groups_1.2M_20-500gens+200rep_nofails.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(df_subMed[20:500,], aes(x = Generation)) +
  geom_line(aes(y = Geometric_mean), color = 'red') +
  ggtitle(paste('Combined demographic trajectories of two groups from Mediterranean region')) + #add a title
  geom_line(data = df_Mal_nofails[20:500,], aes(x = Generation, y = Geometric_mean), color = 'blue') +
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
dev.off()
```
