---
title: "Plot_div"
output: html_document
date: "2024-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/home/newuser/Desktop/Rworks')
```

## Plotting diversity parameters with Martin lab software

```{r}
#filter windows with less sites, may be biased (at the end of each chromosome)
div_in_windows <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.10k.csv',header = T)
div_in_windows_filt <- div_in_windows[div_in_windows$sites == max(div_in_windows$sites),]
colnames(div_in_windows_filt)

#average diversity for each population
pi_Med <- median(div_in_windows_filt$pi_Mediterranean)
pi_Afr <- median(div_in_windows_filt$pi_African)



```

Plotting the results for single chromosome, chr15 and B locus

```{r}
library(ggplot2)

#plot the hig-resolution graph for a single chromosome containing the B locus
png(filename = paste('div.Med_vs_Afr.chr15.10kb+Blocus.png', sep = ""),res = 300,width = 200, height = 160, units = 'mm')
  ggplot(div_in_windows_filt[div_in_windows_filt$scaffold == 'chr15',], aes(x = mid)) +
  geom_line(aes(y = pi_Mediterranean), color = 'red') +
  ggtitle(paste('Diversity in 10k windows, chr15')) + #add a title
  geom_vline(xintercept = 6.3e6, linetype = "dashed", color = "red", linewidth = 0.7) + # add a line to show position of B locus
  geom_line(data = div_in_windows_filt[div_in_windows_filt$scaffold == 'chr15',], aes(x = mid, y = pi_African), color = 'blue') +
  theme(plot.title = element_text(hjust = 0.5)) #adjust the title position

dev.off()
```

### Diversity boxplot for each subgroup

```{r}
library(ggplot2)
library(reshape)

div_100k_sub <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.sub.csv',header = T)
div_100k <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.csv',header = T)


#combine the pi-estimates for a comprehensive boxplot
group_names <- grep('^pi_*',colnames(div_100k_sub),value = T) #find the pattern for diversity of each group name
group_names_pop <- grep('^pi_*',colnames(div_100k),value = T)
div_100k_combine <- cbind(div_100k[,group_names_pop],div_100k_sub[,group_names])

#calculate the median values for each subgroup
pi_medians <- apply(div_100k[,group_names_pop],2,median)
pi_medians_sub <- apply(div_100k_sub[,group_names],2,median)
print(pi_medians)
print(pi_medians_sub)

#boxplot to demonstrate the population structure between groups
div_100k_sub_m <- melt(div_100k_combine)
n <- nrow(div_100k_sub)
div_100k_sub_m$Population <- c(rep('Mediterranean',n),rep('African',n),rep('Mediterranean',7*n),rep('African',2*n))

png(filename = 'Diversity_boxplot_rmoutlier.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_100k_sub_m, aes(x = variable, y = value, colour = Population)) + 
  xlab('Population groups') + ylab('Diversity') +
  geom_boxplot(outlier.shape=NA, notch=FALSE) + #point out the outliers and erase the CIs around the medians
  geom_vline(xintercept = 2.5, linetype = "dashed", color = "red", linewidth = 0.7) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()


#Wilcoxon test between groups to prove they're not statisticaly different
wcx_Med_Afr <- wilcox.test(div_100k_combine$pi_Mediterranean,div_100k_combine$pi_African)
print(wcx_Med_Afr$p.value)
```

### Plot Fst between populations in windows, B locus

```{r}
library(ggplot2)

plot.Fst <- function(chrom,df){
  ymax <- max(df$Fst_Mediterranean_African) + 0.005
  windsize <- max(df$sites)
  div_chrom <- df[df$scaffold %in% chrom,]
  div_chrom <- div_chrom[div_chrom$Fst_Mediterranean_African >= 0,]
  if (length(chrom)>1) {
    chrom <- 'autosome'
  } #if the input is the whole chromosome, generate the corresponding output
  ggplot(div_chrom, aes(x=mid,y=Fst_Mediterranean_African)) + geom_line() + 
    ylim(0,ymax) + #standardise the scale of diversity in the plot
    xlab('Coordinates') + #add a x-axis label 
    geom_vline(xintercept = 6.3e6, linetype = "dashed", color = "red", linewidth = 0.7) + # add a line to show position of B locus
    ggtitle(paste('Fst between Mediterranean and Arican population ',chrom,', window size = ',windsize, sep = "")) + #add a title
    theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
  
}

#plot the high-resolution graph for a single chromosome that contains the B locus
chrom_names <- 'chr15'
png(filename = paste('Div_plots/Fst.autosome.',chrom_names,'.10kb.png', sep = ""),res = 300,width = 200, height = 160, units = 'mm')
  plot.Fst(chrom_names,div_in_windows_filt)
dev.off()


#visualise the window length with middle coordinates
test <- div_in_windows_filt[div_in_windows_filt$scaffold == 'chr15',]
plot(test$mid,test$mid)
```

### Summarise Fst between groups with a boxplot

```{r}
library(ggplot2)
library(reshape)
library(stringr)

div_100k_sub <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.sub.csv', header = T)
div_100k <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.csv',header = T)

Fst_labels <- grep('Fst*',colnames(div_100k_sub),value = T)
div_100k_Fst <- cbind(div_100k_sub[,Fst_labels],div_100k$Fst_Mediterranean_African)
colnames(div_100k_Fst)[ncol(div_100k_Fst)] <- 'Fst_Mediterranean_African'
 
#replace NA with the median for each column
replace_na_with_median <- function(column) {
  # Calculate the median, excluding NA values
  median_value <- median(column, na.rm = TRUE)
  # Replace NA values with the median
  column[is.na(column)] <- median_value
  return(column)
}
div_100k_Fst_num <- as.data.frame(apply(div_100k_Fst, 2,replace_na_with_median))

#calculate the median Fst between groups
Fst_medians <- apply(div_100k_Fst_num, 2, median)
Fst_medians



#group the comparisons accroding to within/between population
# Define the reference list
meta <- read.csv('metadata_detailed.csv')
ref_Med <- unique(meta[meta$Region == 'Mediterranean','Narrow_regions'])
ref_Med[ref_Med == 'Canary Islands'] <- 'Canary'
ref_Afr <- unique(meta[meta$Region == 'African','Narrow_regions'])

#Between group comparisons contain group names from both populations
between_label <- str_detect(colnames(div_100k_Fst_num), str_c(ref_Med, collapse = "|")) & 
  str_detect(colnames(div_100k_Fst_num), str_c(ref_Afr, collapse = "|"))

#reorder the dataframe
col_reorder <- c(colnames(div_100k_Fst_num)[between_label],
                colnames(div_100k_Fst_num[!between_label]))
div_100k_Fst_num_reorder <- div_100k_Fst_num[,col_reorder]

#create input for ggplot
div_100k_Fst_num_m <- melt(div_100k_Fst_num_reorder)
div_100k_Fst_num_m$Comparison <- 'within'
between_label_m <- str_detect(div_100k_Fst_num_m$variable, str_c(ref_Med, collapse = "|")) & 
  str_detect(div_100k_Fst_num_m$variable, str_c(ref_Afr, collapse = "|"))
table(between_label_m)
div_100k_Fst_num_m[between_label_m,'Comparison'] <- 'between'
between_label_t <- str_detect(div_100k_Fst_num_m$variable, str_c('Med', collapse = "|"))
div_100k_Fst_num_m[between_label_t,'Comparison'] <- 'total'
table(div_100k_Fst_num_m$Comparison)

#plot the Fst boxplot
png(filename = 'Fst_boxplot.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_100k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=1, notch=FALSE) + #point out the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()

#remove the outliers for the boxplot
png(filename = 'Fst_boxplot_rmoutlier.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_100k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  ylim(-0.1,0.1) +
  geom_hline(aes(yintercept = 0), col = 'red', alpha = 0.5, lty = 'dashed') +
  geom_boxplot(outlier.shape=NA,notch=FALSE) + #remove the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()

Fst_median_between <- Fst_medians[col_reorder[1:14]]
Fst_median_within <- Fst_medians[col_reorder[15:36]]
t_Fst_median <- t.test(Fst_median_between,Fst_median_within)
```

### Force the negative Fst values to 0 and plot again

```{r}
library(ggplot2)
library(reshape)
library(stringr)

div_100k_sub <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.sub.csv', header = T)
div_100k <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.Call20.100k.csv',header = T)

Fst_labels <- grep('Fst*',colnames(div_100k_sub),value = T)

#Remove the comparisons with small sample size N<3
words_to_remove <- c("Tunisia", "Algeria")
pattern <- paste(words_to_remove, collapse = "|")
filtered_Fst_labels <- Fst_labels[!grepl(pattern, Fst_labels)]

div_100k_Fst <- cbind(div_100k_sub[,filtered_Fst_labels],div_100k$Fst_Mediterranean_African)
colnames(div_100k_Fst)[ncol(div_100k_Fst)] <- 'Fst_Mediterranean_African'
 
#replace negative values with 0
temp <- as.data.frame(lapply(div_100k_Fst, function(x) ifelse(x < 0, 0, x)))

#replace NA with the median for each column
replace_na_with_median <- function(column) {
  # Calculate the median, excluding NA values
  median_value <- median(column, na.rm = TRUE)
  # Replace NA values with the median
  column[is.na(column)] <- median_value
  return(column)
}
div_100k_Fst_num <- as.data.frame(apply(temp, 2,replace_na_with_median))



#calculate the median Fst between groups
Fst_medians <- apply(div_100k_Fst_num, 2, median)
Fst_medians



#group the comparisons accroding to within/between population
# Define the reference list
meta <- read.csv('metadata_detailed.csv')
ref_Med <- unique(meta[meta$Region == 'Mediterranean','Narrow_regions'])
ref_Med[ref_Med == 'Canary Islands'] <- 'Canary'
ref_Afr <- unique(meta[meta$Region == 'African','Narrow_regions'])

#Between group comparisons contain group names from both populations
between_label <- str_detect(colnames(div_100k_Fst_num), str_c(ref_Med, collapse = "|")) & 
  str_detect(colnames(div_100k_Fst_num), str_c(ref_Afr, collapse = "|"))

#reorder the dataframe
col_reorder <- c(colnames(div_100k_Fst_num)[between_label],
                colnames(div_100k_Fst_num[!between_label]))
div_100k_Fst_num_reorder <- div_100k_Fst_num[,col_reorder]

#create input for ggplot
div_100k_Fst_num_m <- melt(div_100k_Fst_num_reorder)
div_100k_Fst_num_m$Comparison <- 'within'
between_label_m <- str_detect(div_100k_Fst_num_m$variable, str_c(ref_Med, collapse = "|")) & 
  str_detect(div_100k_Fst_num_m$variable, str_c(ref_Afr, collapse = "|"))
table(between_label_m)
div_100k_Fst_num_m[between_label_m,'Comparison'] <- 'between'
between_label_t <- str_detect(div_100k_Fst_num_m$variable, str_c('Med', collapse = "|"))
div_100k_Fst_num_m[between_label_t,'Comparison'] <- 'total'
table(div_100k_Fst_num_m$Comparison)

#plot the Fst boxplot
png(filename = 'Fst_boxplot_nonegs_n>=3.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_100k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=1, notch=FALSE) + #point out the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()

#remove the outliers for the boxplot
png(filename = 'Fst_boxplot_rmoutlier_nonegs_n>=3.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_100k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  ylim(0,0.1) +
  geom_hline(aes(yintercept = 0), col = 'red', alpha = 0.5, lty = 'dashed') +
  geom_boxplot(outlier.shape=NA,notch=FALSE) + #remove the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()
```

## Plotting heatmap for divergence between individuals to demonstrate population subdivision

```{r}
dist <- read.table('distmat.eu40.vs.dchry2.3.autosome.4dsites.Call20.dist', header = F, row.names = 1, skip = 1)
meta <- read.csv('metadata_detailed.csv')

colnames(dist) <- row.names(dist)
dist.t <- dist[match(meta$Sample.ID,rownames(dist)),]
dist.t <- dist.t[,match(meta$Sample.ID,colnames(dist))] #group the samples according to their geographic location

dist.m <- as.matrix(dist.t)


library(reshape)


dist.m.melted <- melt(dist.m) #create the input matrix for ggplot heatmap: each row contains rowname, colname and the value for a position in the dxy matrix.
colnames(dist.m.melted) <- c('x','y','dxy')

cl <- c(rep('blue',28),rep('red',12)) #set the colours for labeling geographic groups
```

Plotting the results with higher resolution:

```{r, change savepath before runnning}
library(ggplot2)

png(filename = 'pairwise_dxy_heatmap_autosome_4dsites.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(dist.m.melted, aes(x = x, y = y, fill = dxy)) + #filling the tiles with values in the dxy matrix
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1), axis.text = element_text(size = 6,colour = cl)) + #adjust x-axis label angle and label sizes for both axes 
  geom_tile() + #plot the heatmap
  scale_fill_gradient(low = "white", high = "orange") + #adjust the colors representing values
  coord_fixed() #force tiles to be square shape
dev.off()
```

## Calculate pairwise dxy - geographic distance

```{r}
library(readxl)
meta.deg <- read_xlsx('chrysippus_RVcoll_sequenced_African_added.xlsx')
meta <- read.csv('metadata_detailed.csv')

#Merge metadata with coordinates, then adjust the order of samples to be consistent with metadata
meta.deg2 <- meta.deg[,c('Sample ID',"Lat (Dec. Deg)","Long (Dec. Deg)")]
meta.all <- merge(meta,meta.deg2,by.x = 'Sample.ID',by.y = 'Sample ID')
meta.all <- meta.all[match(meta$Sample.ID,meta.all$Sample.ID),]

column_names <- colnames(meta.all)
column_names[c(4,5)] <- c('Lat','Long')
colnames(meta.all) <- column_names

#Take the mean coordinates for missing ones among the Ghanean samples
#Before data input, missing coordiantes for the Catalonian sample are susubstituted with the position of collection on Google map
val1 <- meta.all[meta.all$Narrow_regions == 'Ghana','Lat'] 
val2 <- meta.all[meta.all$Narrow_regions == 'Ghana','Long']
meta.all[meta.all$Sample.ID == c('IG19G105','IG19G114'),'Lat'] <- mean(val1[!is.na(val1)])
meta.all[meta.all$Sample.ID == c('IG19G105','IG19G114'),'Long'] <- mean(val2[!is.na(val2)])

#One possibe way to calculate distance from coordinates, from Thomas
deg2rad <- function(deg) {(deg * pi) / (180)}
coord2dist <- function(lat, lon, lat_1, lon_1) {
  R <- 6371
  a = (sin((deg2rad(lat_1) - deg2rad(lat)) / 2) ** 2) + cos(deg2rad(lat)) * cos(deg2rad(lat_1)) * (sin((deg2rad(lon_1) - deg2rad(lon)) / 2) ** 2)
  d = 2*R*atan2(sqrt(a), sqrt(1 - a))
  return(d)
  
}

#Create distance matrix 
div_vs_dist <- matrix(0,nrow = 1600, ncol = 3)
div_vs_dist <- as.data.frame(div_vs_dist)
colnames(div_vs_dist) <- c('x','y','dist')

for (i in 1:40) {
  for (j in 1:40) {
    div_vs_dist[(i-1)*40+j,'y'] <- meta.all[i,'Sample.ID']
    div_vs_dist[(i-1)*40+j,'x'] <- meta.all[j,'Sample.ID'] #fill in x first, then fill in y
    div_vs_dist[(i-1)*40+j,'dist'] <- coord2dist(meta.all[i,'Lat'],meta.all[i,'Long'],meta.all[j,'Lat'],meta.all[j,'Long'])
  }
}
```

*Run the code block to input divergence matrix before running this block*

```{r}
library(ggplot2)
library(reshape)

dist <- read.table('distmat.eu40.vs.dchry2.3.autosome.4dsites.Call20.dist', header = F, row.names = 1, skip = 1)
meta <- read.csv('metadata_detailed.csv')

colnames(dist) <- row.names(dist)
dist.t <- dist[match(meta$Sample.ID,rownames(dist)),]
dist.t <- dist.t[,match(meta$Sample.ID,colnames(dist))] #group the samples according to their geographic location

dist.m <- as.matrix(dist.t)
dist.m.melted <- melt(dist.m) #create the input matrix for ggplot heatmap: each row contains rowname, colname and the value for a position in the dxy matrix.
colnames(dist.m.melted) <- c('x','y','dxy')

cl <- c(rep('blue',28),rep('red',12)) #set the colours for labeling geographic groups

#check the consistency between divergence matrix and geodistance matrix, then imput divergence values
sum(div_vs_dist$x == dist.m.melted$x)
sum(div_vs_dist$y == dist.m.melted$y)
div_vs_dist$dxy <- dist.m.melted$dxy

# plot divergence against distance, exclude self-comparsions
between_label <- which(div_vs_dist$x != div_vs_dist$y)
div_vs_dist_between <- div_vs_dist[between_label,]

#label the pairs of samples from the same location, exclude them from the plot
div_vs_dist_between$Population <- 0
for (i in 1:nrow(div_vs_dist_between)) {
  sample.x <- div_vs_dist_between[i,'x']
  sample.y <- div_vs_dist_between[i,'y']
  if(meta.all[meta.all$Sample.ID == sample.x,'Narrow_regions'] == meta.all[meta.all$Sample.ID ==             sample.y,'Narrow_regions']){
    div_vs_dist_between[i,'Population'] <- 1
  }
}
div_vs_dist_between_regions <- div_vs_dist_between[div_vs_dist_between$Population == 0,]

#group the pairs of samples with: between/within population
div_vs_dist_between_regions$Population <- 'between'
for (i in 1:nrow(div_vs_dist_between_regions)) {
  sample.x <- div_vs_dist_between_regions[i,'x']
  sample.y <- div_vs_dist_between_regions[i,'y']
  if(meta.all[meta.all$Sample.ID == sample.x,'Region'] == meta.all[meta.all$Sample.ID ==             sample.y,'Region']){
    div_vs_dist_between_regions[i,'Population'] <- 'within' #label purple if pair of samples from the same population
  }
}


#function for adding regression line to ggplot
lm_eqn <- function(df){
    m <- lm(dxy ~ dist, data = df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
} #function for adding legend for the regression line

# output the graph
png(filename = 'Dvg_vs_dist.autosome.4d.png',res = 300,width = 200, height = 160, units = 'mm')

ggplot(data = div_vs_dist_between_regions, aes(x = dist, y = dxy, color = Population)) +
    geom_smooth(method = "lm", se=FALSE, color="red", formula = y ~ x, lty = 'dashed') + #add regression line on the plot
    geom_point() +
    geom_text(x = 1000, y = 0.046, label = lm_eqn(div_vs_dist_between_regions), parse = TRUE, cex = 4, col = 'red') #adjust the position and size of the regression line legend

dev.off()

m <- lm(dxy ~ dist, data = div_vs_dist_between_regions)
summary(m)
```

Plot net divergence against distance: [dxy is smaller than population diversity, may not be applicable]{.underline}

### Plot the dxy vs. distance graph for the Mediterranean group alone

```{r}
#find the pairs of samples from Mediterranean region
for (i in 1:nrow(div_vs_dist_between_regions)) {
  sample.x <- div_vs_dist_between_regions[i,'x']
  sample.y <- div_vs_dist_between_regions[i,'y']
  if(meta.all[meta.all$Sample.ID == sample.x,'Region'] ==  'Mediterranean' && meta.all[meta.all$Sample.ID == sample.y,'Region'] ==  'Mediterranean'){
    div_vs_dist_between_regions[i,'Population'] <- 'withinMed' #label purple if pair of samples from the same population
  }
}
div_vs_dist_between_Med <- div_vs_dist_between_regions[div_vs_dist_between_regions$Population == 'withinMed',]

#function for adding regression line to ggplot
lm_eqn <- function(df){
    m <- lm(dxy ~ dist, data = df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
} #function for adding legend for the regression line

# output the graph
png(filename = 'Dvg_vs_dist.autosome.4d.Med.png',res = 300,width = 200, height = 160, units = 'mm')

ggplot(data = div_vs_dist_between_Med, aes(x = dist, y = dxy, color = Population)) +
    geom_smooth(method = "lm", se=FALSE, color="red", formula = y ~ x, lty = 'dashed') + #add regression line on the plot
    geom_point() +
    geom_text(x = 770, y = 0.0454, label = lm_eqn(div_vs_dist_between_Med), parse = TRUE, cex = 4, col = 'red') #adjust the position and size of the regression line legend

dev.off()

m <- lm(dxy ~ dist, data = div_vs_dist_between_Med)
summary(m)
```

## Plot the putative neutral site diversity between populations (4d-sites) within windows

```{r}
#filter windows with less sites, may be biased (at the end of each chromosome)
div_4d <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.4dsites.Call20.10k.csv',header = T)
div_4d_filt <- div_4d[div_4d$sites == max(div_4d$sites),]


#plot 4d-diversity per chromosome for each population
library(ggplot2)

plot.div.Med <- function(chrom,df){
  ymax <- max(df$pi_Mediterranean) + 0.005
  windsize <- max(df$sites)
  div_chrom <- df[df$scaffold %in% chrom,]
  if (length(chrom)>1) {
    chrom <- 'autosome'
  }
  ggplot(div_chrom, aes(x=1:nrow(div_chrom),y=pi_Mediterranean)) + geom_line() + 
    ylim(0,ymax) + #standardise the scale of diversity in the plot
    xlab('Windows') + #add a x-axis label 
    ggtitle(paste('Diversity of',chrom,', window size =',windsize)) + #add a title
    theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
  
}

plot.div.Afr <- function(chrom,df){
  ymax <- max(df$pi_African) + 0.005
  windsize <- max(df$sites)
  div_chrom <- df[df$scaffold %in% chrom,]
  if (length(chrom)>1) {
    chrom <- 'autosome'
  }
  ggplot(div_chrom, aes(x=1:nrow(div_chrom),y=pi_African)) + geom_line() + 
    ylim(0,ymax) + #standardise the scale of diversity in the plot
    xlab('Windows') + #add a x-axis label 
    ggtitle(paste('Diversity of',chrom,', window size =',windsize)) + #add a title
    theme(plot.title = element_text(hjust = 0.5)) #adjust the title position
  
}

plot.div.Med(unique(div_4d_filt$scaffold),div_4d_filt)
plot.div.Afr(unique(div_4d_filt$scaffold),div_4d_filt)


chrom_names <- unique(div_in_windows_filt$scaffold)
for (i in 1:length(chrom_names)) {

png(filename = paste('Div_plots/div.Med.autosome.',chrom_names[i],'.10kb.png', sep = ""),res = 300,width = 200, height = 160, units = 'mm')
  plot.div.Med(chrom_names[i],div_in_windows_filt)
dev.off()

}
```

### Boxplot for 4d-site diversity for each sub-group

```{r}
library(ggplot2)
library(reshape)

div_4d_10k_sub <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.4dsites.Call20.10k(subgroups).csv')
div_4d_10k <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.4dsites.Call20.10k.csv',header = T)


#combine the pi-estimates for a comprehensive boxplot
group_names <- grep('^pi_*',colnames(div_4d_10k_sub),value = T) #find the pattern for diversity of each group name
group_names_pop <- grep('^pi_*',colnames(div_4d_10k),value = T)
div_4d_10k_combine <- cbind(div_4d_10k[,group_names_pop],div_4d_10k_sub[,group_names])

#calculate the median values for each subgroup
pi_medians_4d <- apply(div_4d_10k[,group_names_pop],2,median)
pi_medians_4d_sub <- apply(div_4d_10k_sub[,group_names],2,median)
print(pi_medians_4d)
print(pi_medians_4d_sub)

#boxplot to demonstrate the population structure between groups
div_4d_10k_sub_m <- melt(div_4d_10k_combine)
n <- nrow(div_4d_10k_sub)
div_4d_10k_sub_m$Population <- c(rep('Mediterranean',n),rep('African',n),rep('Mediterranean',7*n),rep('African',2*n))

png(filename = 'Diversity_boxplot_4d_rmoutlier.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_4d_10k_sub_m, aes(x = variable, y = value, colour = Population)) + 
  xlab('Population groups') + ylab('4d-Diversity') +
  geom_boxplot(outlier.shape=NA, notch=FALSE) + #point out the outliers and erase the CIs around the medians
  geom_vline(xintercept = 2.5, linetype = "dashed", color = "red", linewidth = 0.7) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()


#Wilcoxon test between groups to prove they're not statisticaly different
wcx_Med_Afr <- wilcox.test(div_4d_10k_combine$pi_Mediterranean,div_4d_10k_combine$pi_African)
print(wcx_Med_Afr$p.value)
```

### Plot Fst between groups based on the 4d-sites

```{r}
library(ggplot2)
library(reshape)
library(stringr)

div_4d_10k_sub <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.4dsites.Call20.10k(subgroups).csv')
div_4d_10k <- read.csv('div.eu40.vs.dchry2.3.autosome.DP8_100.GQ20.4dsites.Call20.10k.csv',header = T)

Fst_labels <- grep('Fst*',colnames(div_4d_10k_sub),value = T)
div_4d_10k_Fst <- cbind(div_4d_10k_sub[,Fst_labels],div_4d_10k$Fst_Mediterranean_African)
colnames(div_4d_10k_Fst)[ncol(div_4d_10k_Fst)] <- 'Fst_Mediterranean_African'
 
#replace NA with the median for each column
replace_na_with_median <- function(column) {
  # Calculate the median, excluding NA values
  median_value <- median(column, na.rm = TRUE)
  # Replace NA values with the median
  column[is.na(column)] <- median_value
  return(column)
}
div_4d_10k_Fst_num <- as.data.frame(apply(div_4d_10k_Fst, 2,replace_na_with_median))

#calculate the median Fst between groups
Fst_medians <- apply(div_4d_10k_Fst_num, 2, median)
Fst_medians



#group the comparisons accroding to within/between population
# Define the reference list
meta <- read.csv('metadata_detailed.csv')
ref_Med <- unique(meta[meta$Region == 'Mediterranean','Narrow_regions'])
ref_Med[ref_Med == 'Canary Islands'] <- 'Canary'
ref_Afr <- unique(meta[meta$Region == 'African','Narrow_regions'])

#Between group comparisons contain group names from both populations
between_label <- str_detect(colnames(div_4d_10k_Fst_num), str_c(ref_Med, collapse = "|")) & 
  str_detect(colnames(div_4d_10k_Fst_num), str_c(ref_Afr, collapse = "|"))

#reorder the dataframe
col_reorder <- c(colnames(div_4d_10k_Fst_num)[between_label],
                colnames(div_4d_10k_Fst_num[!between_label]))
div_4d_10k_Fst_num_reorder <- div_4d_10k_Fst_num[,col_reorder]

#create input for ggplot
div_4d_10k_Fst_num_m <- melt(div_4d_10k_Fst_num_reorder)
div_4d_10k_Fst_num_m$Comparison <- 'within'
between_label_m <- str_detect(div_4d_10k_Fst_num_m$variable, str_c(ref_Med, collapse = "|")) & 
  str_detect(div_4d_10k_Fst_num_m$variable, str_c(ref_Afr, collapse = "|"))
table(between_label_m)
div_4d_10k_Fst_num_m[between_label_m,'Comparison'] <- 'between'
between_label_t <- str_detect(div_4d_10k_Fst_num_m$variable, str_c('Med', collapse = "|"))
div_4d_10k_Fst_num_m[between_label_t,'Comparison'] <- 'total'
table(div_4d_10k_Fst_num_m$Comparison)

#plot the Fst boxplot
png(filename = 'Fst_boxplot_4d.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_4d_10k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  geom_boxplot(outlier.colour="black", outlier.shape=16, outlier.size=1, notch=FALSE) + #point out the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()

#remove the outliers for the boxplot
png(filename = 'Fst_boxplot_4d_rmoutlier.png',res = 300,width = 200, height = 160, units = 'mm')
ggplot(div_4d_10k_Fst_num_m, aes(x = variable, y = value, color = Comparison)) + 
  xlab('Group comparisons') + ylab('Fst') +
  ylim(-0.1,0.1) +
  geom_hline(aes(yintercept = 0), col = 'red', alpha = 0.5, lty = 'dashed') +
  geom_boxplot(outlier.shape=NA,notch=FALSE) + #remove the outliers and erase the CIs around the medians
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) #rotate the x-axis label direction
dev.off()
```
